Using SMT engine to generate Symbolic Automata - Extended version\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-02-02 2017-Xudong-SMT/FM-2018-extended-RR.tex
First elements for elimination

\title{Using SMT engine to generate Symbolic Automata\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Xudong Qin\inst{1,2}  \ \ \  Simon Bliudze\inst{3} \ \ \ Eric Madelaine\inst{1}
  \ \ \  Min Zhang\inst{2}} 
\institute{Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia Antipolis, France
        \and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China
\and INRIA Lille -- Nord Europe, 40 avenue Halley,
59650 Villeneuve d'Ascq,
France}
\date{}                                           % Activate to display a given date or no date

@@

For example, in the running example, in the raw output listing in
Appendix~ref{}, in the first OT, you find variables: 
\begin{itemize}
   \item $\OTvar{:hb\_B:13:1}$ that is the behaviour of hole ``B''
   \item $\OTvar{b1:sva\_SV0:1:1}$ variable ``b1'' from vector ``SV0''
   \item $\OTvar{:ra:1:1}$ result action of the current OT
\end{itemize}

@@

In this example, there are 327 satisfiable predicates for the total of 97920 candidate open transitions, computed in 105384 milliseconds (~1.45 minutes). 
The computation was carried out on an Intel Core i7 at 2.5GHz with 16GB of RAM.

@@

\subsection{Simplification of Predicates}
\TODO{I keep it here for now, but I think we shall not include it in
  the FM submission, this section is not mature enough}

@@

Before:
      \begin{mathpar}
    \inferrule*[myfraction=\reddottedrule]
    {\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
      \{\xrightarrow{hb}\},\\
\qquad b1=true \land b2=true \land hb=fail(b0) \land b1=b2 \land (b1\cup b2
\implies b0) \land v=\underline{fail}, \{t := 4\}
        }
    {<s0,t0> \xrightarrow {v} <s1,t1>}
    \end{mathpar}

@@

\QIN{
We want to reason about the equations to eliminate the redundancy. 
Considering pNet has a special tree-structure, the replacement between equations should be one-way, the direction is from leaves to the top level, until the root pNet node. 
And it is easy to figure out the intermediate variables by the constraint rules.
While it is not needed figure out if an equation is "true" , it can be left to Z3.
}

@@

\paragraph{Term Rewriting Rules}
Predicates already have a format declaring what it verifies. According to that previous work, we define the predicates.
Let $\mylangle\overline{\pNet},\overline{S},\symb{SV}_k^{k\in K} \myrangle$
be a pNet. Choose a synchronisation vector $SV=clone(SV_k) ={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v' \, ,G_k $, for $k\in K$. The
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions.
\begin{definition}[Predicate]
\[\Pred(SV, \overline{\pNet},\overline{S}, v) = 
\begin{array}{l}
\forall {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.
\\
v_m=a'_i\land \, b_j=b'_j \land v=v' \land G_k
\end{array}\]
\end{definition}
\QIN{
The new predicates introduce more hypotheses for term rewriting. However, not every variable can be substituted. 
Here we introduce several rewriting rules.
}

\begin{align*}
   &v_m=a'_i \land v_m=v'_m \rightarrow a'_i = v'_m &[RL-1]\\\\
   &v_m=a'_i \land v=v' \rightarrow v_m=v ~~ \texttt{if} ~~ a'_i = v', ~~ a'_i,v' \in InterVar &[RL-2]\\\\
   &v_m=foo'_i(a_1,...,a_p,...,a_n) \land v=v' \rightarrow v_m=foo'_i(a_1,...,v,...,a_n) \\ 
   &\hspace{15em}\texttt{if} ~~ a_p = v', ~~ a_p,v \in InterVar &[RL-3]\\\\
   &b_j=b'_j \land v=v' \rightarrow b_j=v ~~ \texttt{if} ~~ b'_j = v', ~~ b'_j,v' \in InterVar &[RL-4]\\\\
   &b_j=foo'_i(a_1,...,a_p,...,a_n) \land v=v' \rightarrow b_j=foo'_i(a_1,...,v,...,a_n) \\ 
   &\hspace{15em}\texttt{if} ~~ a_p = v', ~~ a_p,v \in InterVar &[RL-5]\\\\
\end{align*}

\QIN{
Rule RL-1 applies on when the result action of subnets matching other arguments of the chosen SV. It merges the predicates from the subnets with the new generated predicates.
Rule RL-2 shows that if the result of the SV is also in its arguments, we can straightly get an equation between the subnet action and  the result action. RL-3 is the situation that there is an expression instead of a variable contains an intermediate argument. The substitution will only occur on the argument.
Actually, what happened is similar if there is a hole behaviour instead of a subnet result action attending in the equation. So we get RL-4 and RL-5. However, every hole behaviour should be kept through out the computing. That's why there no rule similar to the RL-1 for hole behaviours.
}

>>>>

\title{Using SMT engine to generate Symbolic Automata - Extended version\thanks{This work was partially 
funded by the Associated Team FM4CPS
  between INRIA and ECNU, Shanghai}}
\author{ Xudong Qin\inst{1,2}  \ \ \  Simon Bliudze\inst{3} \ \ \ Eric Madelaine\inst{1}
  \ \ \  Min Zhang\inst{2}} 
\institute{Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia Antipolis, France
        \and Shanghai Key Laboratory of Trustworthy Computing, ECNU, China
\and INRIA Lille -- Nord Europe, 40 avenue Halley,
59650 Villeneuve d'Ascq,
France}
\date{}                                           % Activate to display a given date or no date

@@

For example, in the running example, in the raw output listing in
Appendix~\ref{appendix:fullOutput}, in the first OT, you find variables: 
\begin{itemize}
   \item $\OTvar{:hb\_B:13:1}$ that is the behaviour of hole ``B''
   \item $\OTvar{b1:sva\_SV0:1:1}$ variable ``b1'' from vector ``SV0''
   \item $\OTvar{:ra:1:1}$ result action of the current OT
\end{itemize}

@@

In this example, there are 327 satisfiable predicates for the total of 97920 candidate open transitions, computed in 105384 milliseconds (\textasciitilde{}1.45 minutes). 
The computation was carried out on an Intel Core i7 at 2.5GHz with 16GB of RAM.

@@

%\end{document}

\subsection{Simplification of Predicates}

@@

Before:
\ERIC{Also this OT is not the real original one (as in RoughResult.log), pieces have already been simplied...}

      \begin{mathpar}
    \inferrule*[myfraction=\reddottedrule]
    {\{s0~{\xrightarrow{fail(true)}} ~s1, t0~{\xrightarrow{start(true)}} ~t1\}
      \{\xrightarrow{hb}\},\\
\qquad b1=true \land b2=true \land hb=fail(b0) \land b1=b2 \land (b1\cup b2
\implies b0) \land v=\underline{fail}, \{t := 4\}
        }
    {<s0,t0> \xrightarrow {v} <s1,t1>}
    \end{mathpar}

@@

\TODO{

% We want to reason about the equations to eliminate the redundancy. 
% Considering pNet has a special tree-structure, the replacement between equations should be one-way, the direction is from leaves to the top level, until the root pNet node. 
And it is easy to figure out the intermediate variables by the constraint rules.
% While it is not needed figure out if an equation is "true" , it can be left to Z3.

@@

%% \paragraph{Term Rewriting Rules}
%% Predicates already have a format declaring what it verifies. According to that previous work, we define the predicates.
%% Let $\mylangle\overline{\pNet},\overline{S},\symb{SV}_k^{k\in K} \myrangle$
%% be a pNet. Choose a synchronisation vector $SV=clone(SV_k) ={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v' \, ,G_k $, for $k\in K$. The
%% predicate $\Pred$ relating
%% the actions of the involved sub-pNets and the resulting actions.
%% \begin{definition}[Predicate]
%% \[\Pred(SV, \overline{\pNet},\overline{S}, v) = 
%% \begin{array}{l}
%% \forall {(a'_i)}^{i\in I},
%% {(b'_j)}^{j\in J},v'.
%% \\
%% v_m=a'_i\land \, b_j=b'_j \land v=v' \land G_k
%% \end{array}\]
%% \end{definition}
%% \QIN{
%% The new predicates introduce more hypotheses for term rewriting. However, not every variable can be substituted. 
%% Here we introduce several rewriting rules.
%% }

%% \begin{align*}
%%    &v_m=a'_i \land v_m=v'_m \rightarrow a'_i = v'_m &[RL-1]\\\\
%%    &v_m=a'_i \land v=v' \rightarrow v_m=v ~~ \texttt{if} ~~ a'_i = v', ~~ a'_i,v' \in InterVar &[RL-2]\\\\
%%    &v_m=foo'_i(a_1,...,a_p,...,a_n) \land v=v' \rightarrow v_m=foo'_i(a_1,...,v,...,a_n) \\ 
%%    &\hspace{15em}\texttt{if} ~~ a_p = v', ~~ a_p,v \in InterVar &[RL-3]\\\\
%%    &b_j=b'_j \land v=v' \rightarrow b_j=v ~~ \texttt{if} ~~ b'_j = v', ~~ b'_j,v' \in InterVar &[RL-4]\\\\
%%    &b_j=foo'_i(a_1,...,a_p,...,a_n) \land v=v' \rightarrow b_j=foo'_i(a_1,...,v,...,a_n) \\ 
%%    &\hspace{15em}\texttt{if} ~~ a_p = v', ~~ a_p,v \in InterVar &[RL-5]\\\\
%% \end{align*}

%% \QIN{
%% Rule RL-1 applies on when the result action of subnets matching other arguments of the chosen SV. It merges the predicates from the subnets with the new generated predicates.
%% Rule RL-2 shows that if the result of the SV is also in its arguments, we can straightly get an equation between the subnet action and  the result action. RL-3 is the situation that there is an expression instead of a variable contains an intermediate argument. The substitution will only occur on the argument.
%% Actually, what happened is similar if there is a hole behaviour instead of a subnet result action attending in the equation. So we get RL-4 and RL-5. However, every hole behaviour should be kept through out the computing. That's why there no rule similar to the RL-1 for hole behaviours.
%% }
